---
title: "Karyotype diagrams for H3K27me3 CUT&RUN data (Greer Lab)"
author: "Shannan Ho Sui"
date: "1/12/2022"
output:
   html_document:
      code_folding: hide
      highlights: pygments
      theme: default
      toc: true
      toc_float:
         collapsed: true
         smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

Using ggbio to plot distribution of H3K27me3 signal across chromosomes.


```{r Functions}
library(GenomicRanges)
library(ggbio)
library(rtracklayer)
library(diffloop)
library(viridis)

setwd("~/Dropbox (Harvard University)/HBC Team Folder (1)/Consults/eric_greer/cutnrun_analysis_of_microprotein_hbc04238/ggbio/")

## Function to load broadPeak files into gRanges object
load_broadPeak <- function(file_broadPeak) {
  # specify extra columns for bed format
  extraCols_broadPeak <- c(signalValue = "numeric", pValue = "numeric", qValue = "numeric")
  
  # import as bed file
  gr_broadPeak <- import.bed(file_broadPeak, 
                             extraCols = extraCols_broadPeak)
  # remove nonstandard chromosomes
  gr_broadPeak <- keepStandardChromosomes(gr_broadPeak, pruning.mode = "tidy")
 
  # add "chr" to the beginning of the seqnames using the diffloop package
  gr_broadPeak <- addchr(gr_broadPeak)
  return(gr_broadPeak)
} 

## Function to plot karyogram

plot_custom_karyogram <- function(gr_broadPeak) {
  # load hg19 annotation package for karyogram. Note that this is not updated for hg38
  data(ideoCyto, package = "biovizBase")
  
  # add seqlengths and reorder them
  seqlengths(gr_broadPeak)
  seqlengths(gr_broadPeak) <- seqlengths(ideoCyto$hg19)[names(seqlengths(gr_broadPeak))]
  gr_broadPeak <- keepSeqlevels(gr_broadPeak, paste0("chr", c(1:22, "X")))
  
  # plot karyogram using ggbio; basic layout
  p <- autoplot(gr_broadPeak, layout = "karyogram", aes(color=log(score)))
  return(p)
}
```

## Plot each sample in a karyogram

This first section tests out the method to create draft figures that will be iterated on.

```{r plotBySample}

file_broadPeak <- "../data/macs2/X2020-10-27-CUTRUN-KT-MB231-12-OE1-H3K27me3-1_S12_peaks.broadPeak"
S12 <- load_broadPeak(file_broadPeak)

# plot and write to pdf
#pdf("OE1-H3K27me3-1_S12_v2.pdf")
plot_custom_karyogram(S12) + labs(title = "OE1-H3K27me3-1_S12") + 
  scale_color_viridis(limits = c(2,6))
#dev.off()

file_broadPeak <- "../data/macs2/X2020-10-27-CUTRUN-KT-MB231-13-OE1-H3K27me3-2_S13_peaks.broadPeak"
S13 <- load_broadPeak(file_broadPeak)

#pdf("OE1-H3K27me3-2_S13_v2.pdf")
plot_custom_karyogram(S13) + labs(title = "OE1-H3K27me3-2_S13") + 
  scale_color_viridis(limits = c(2,6))
#dev.off()

file_broadPeak <- "../data/macs2/X2020-10-27-CUTRUN-KT-MB231-04-EV-H3K27me3-1_S4_peaks.broadPeak"
S4 <- load_broadPeak(file_broadPeak)

#pdf("EV-H3K27me3-1_S4_v2.pdf")
plot_custom_karyogram(S4) + labs(title = "EV-H3K27me3-1_S4") + 
  scale_color_viridis(limits = c(2,6))
#dev.off()

```

## Now plot merged replicates

Replicates for each condition were merged using bedops. For each overlapping region, the maximum score was used as the enrichment value.

```{r plotMergedReps}

# Plot OE
#file_broadPeak <- "OE.merge.bed"
file_broadPeak <- "OE.merge_S12_S13_S16_S17_S18.bed"
gr_broadPeak_OE <- load_broadPeak(file_broadPeak)

# plot summary statistics to check which scale to use for the color heatmap
summary(log(gr_broadPeak_OE$score))
gr_broadPeak_OE[log(gr_broadPeak_OE$score) > 5]

pdf("OE_intersect.pdf")
plot_custom_karyogram(gr_broadPeak_OE) + 
  scale_color_viridis(limits = c(2,6)) +
  labs(title = "OE merged")
dev.off()

###################

# Plot EV
#file_broadPeak <- "EV.merge.bed"
file_broadPeak <- "EV.merge_S4_S5_S6.bed"

gr_broadPeak_EV <- load_broadPeak(file_broadPeak)

# plot summary statistics to check which scale to use for the color heatmap
summary(log(gr_broadPeak_EV$score))

pdf("EV_intersect.pdf")
plot_custom_karyogram(gr_broadPeak_EV) + 
  scale_color_viridis(limits = c(2,6)) +
  labs(title = "EV merged")
dev.off()

########################3

## Plot NG
#file_broadPeak <- "NG.merge.bed"
file_broadPeak <- "NG.merge_S3_S8_S9.bed"
gr_broadPeak_NG <- load_broadPeak(file_broadPeak)

# plot summary statistics to check which scale to use for the color heatmap
summary(log(gr_broadPeak_NG$score))

pdf("NG_intersect.pdf")
plot_custom_karyogram(gr_broadPeak_NG) +
  scale_color_viridis(limits = c(2,6)) +
  labs(title = "ng merged")
dev.off()

########################3

## Plot KO
#file_broadPeak <- "KO.merge.bed"
file_broadPeak <- "KO.merge_S22_S23_S24.bed"
gr_broadPeak_KO <- load_broadPeak(file_broadPeak)

summary(log(gr_broadPeak_KO$score))

pdf("KO_intersect.pdf")
plot_custom_karyogram(gr_broadPeak_KO) +
  scale_color_viridis(limits = c(2,6)) +
  labs(title = "KO merged")
dev.off()


```

## Now plot the foldchange between conditions

### EV versus OE

I used the same directionality that Meeta used in the Diffbind analysis. That is, negative "scores" are the negative foldchanges when comparing EV to OE, meaning that a negative value indicates that H3K27me3 marks gained in OE.


```{r fc_plot_EV_vs_OE}

# Load diffbind data
OE_gained_file <- "../diffbind_analysis/results/H3k27me3_mark_gained_in_OE_sig.txt"
OE_lost_file <- "../diffbind_analysis/results/H3k27me3_mark_lost_in_OE_sig.txt"

OE_gained <- read.delim(OE_gained_file)
OE_lost <- read.delim(OE_lost_file)

# concatenate positive and negative foldchanges for EV vs. OE comparison
EV_vs_OE <- rbind(OE_gained, OE_lost)
name <- paste("chr", EV_vs_OE$seqnames, EV_vs_OE$start, sep="_")

# set foldchange as the score
EV_vs_OE <- EV_vs_OE[,c("seqnames", "start", "end","width", "Fold" )]
EV_vs_OE$width <- name
colnames(EV_vs_OE) <- c("chr", "start", "end", "name", "score")

# create gRanges object from dataframe and clean up
gr_EV_vs_OE <- makeGRangesFromDataFrame(EV_vs_OE)
mcols(gr_EV_vs_OE)$score <- EV_vs_OE$score

gr_EV_vs_OE <- keepStandardChromosomes(gr_EV_vs_OE, pruning.mode = "tidy")
gr_EV_vs_OE <- addchr(gr_EV_vs_OE)

# plot summary statistics to check which scale to use for the color heatmap
summary(gr_EV_vs_OE$score)

#pdf("EV_vs_OE_fc.pdf")
plot_custom_karyogram(gr_EV_vs_OE) +
  scale_color_viridis(limits = c(-5,6)) +
  labs(title = "EV vs. OE (foldchange)")
#dev.off()

```

### ng VS KO

Similarly for ng_vs_KO, negative values represent increased H3K27me3 in KO.

```{r fc_plot_ng_vs_KO}

# Load diffbind data
KO_gained_file <- "../diffbind_analysis/results/H3k27me3_mark_gained_in_KO_sig.txt"
KO_lost_file <- "../diffbind_analysis/results/H3k27me3_mark_lost_in_KO_sig.txt"

KO_gained <- read.delim(KO_gained_file)
KO_lost <- read.delim(KO_lost_file)

# concatenate positive and negative foldchanges for ng vs. KO comparison
ng_vs_KO <- rbind(KO_gained, KO_lost)
name <- paste("chr", ng_vs_KO$seqnames,ng_vs_KO$start, sep="_")

# set foldchange as the score
ng_vs_KO <- ng_vs_KO[,c("seqnames", "start", "end","width", "Fold" )]
ng_vs_KO$width <- name
colnames(ng_vs_KO) <- c("chr", "start", "end", "name", "score")

# create gRanges object from dataframe and clean up
gr_ng_vs_KO <- makeGRangesFromDataFrame(ng_vs_KO)
mcols(gr_ng_vs_KO)$score <- ng_vs_KO$score

gr_ng_vs_KO <- keepStandardChromosomes(gr_ng_vs_KO, pruning.mode = "tidy")
gr_ng_vs_KO <- addchr(gr_ng_vs_KO)

# plot summary statistics to check which scale to use for the color heatmap
summary(gr_ng_vs_KO$score)

#pdf("ng_vs_KO_fc.pdf")
plot_custom_karyogram(gr_ng_vs_KO) +
  scale_color_viridis(limits = c(-5,6)) +
  labs(title = "ng vs. KO (foldchange)")
#dev.off()

```
